\pkg{systemfit}: A Package for Estimating Systems of Simultaneous Equations in \proglang{R}

### 2008-09-12 paper/trunk/systemfit.Rnw
moved appendix section "Computation times" (systemfit_timings.Snw) into master file (systemfit.Rnw)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@959 591db2bd-e0a0-4d22-8c19-2192dbb7da31

\input{systemfit_timings.tex}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Computation times}\label{sec:timings}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\SweaveOpts{prefix.string=trash/timings}
<<echo=FALSE>>=
options( prompt = "R> " )
@

Theoretically, one would expect
that the calculations with the \pkg{Matrix} package are
faster and more robust than calculations with the traditional method.
To test this hypothesis, we use function \code{createSystemfitModel}
to create a medium-sized multi-equation model with 8~equations,
10~regressors in each equation (without constant),
and 750~observations.
Then, we estimated this model with and without using the \pkg{Matrix} package.
Finally, the results are compared.
<<systemfit_timings>>=
library( "systemfit" )
set.seed( 1 )
systemfitModel <- createSystemfitModel( nEq = 8, nReg = 10, nObs = 750 )
system.time(
   fitMatrix <- systemfit( systemfitModel$formula, method = "SUR",
      data = systemfitModel$data )
)
system.time(
   fitTrad <- systemfit( systemfitModel$formula, method = "SUR",
      data = systemfitModel$data, useMatrix = FALSE )
)
all.equal( fitMatrix, fitTrad )
@
The returned computation times clearly show
that using the \pkg{Matrix} package makes the estimation faster.
The comparison of the estimation results shows
that both methods return the same results.
The only differences between the returned objects
are --- as expected --- the \code{call}
and the stored control variable \code{useMatrix}.

However, the estimation of rather small models is much slower
with the \pkg{Matrix} package than without this package.
Moreover, the differences in computation time accumulate,
if the estimation is iterated.
<<systemfit_timings>>=
smallModel <- createSystemfitModel( nEq = 3, nReg = 4, nObs = 50 )
system.time(
   fitSmallMatrix <- systemfit( smallModel$formula, method = "SUR",
      data = smallModel$data, maxit = 500 )
)
system.time(
   fitSmallTrad <- systemfit( smallModel$formula, method = "SUR",
      data = smallModel$data, maxit = 500, useMatrix = FALSE )
)
all.equal( fitSmallMatrix, fitSmallTrad )
@
As mentioned above, the usage of the \pkg{Matrix} package
clearly increases the computation times
for iterated (SUR) estimations of small models with small data sets.
